### 2021-10-15 plos_latex_template.tex
Rocío: edits to rule 4

A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, races and ethniticities, socioeconomic statuses, and other aspects of diversity. 
Go beyond balancing all genders in this effort, and pay attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}). To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. If you already have an organizing team, check for gaps in its composition.

% Rocío: what about ethnicity instead of race? %i am not sure

% Rocío: Idea of this paragraph: build a diverse team

Gathering a diverse team will only work if there is real inclusion. People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity. This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups [cite design justice]. A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear.

% Rocío: Idea of this paragraph: real inclusion when having diverse members

Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams, but the positive outcomes are worthwhile. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging (see \textbf{Rule \ref{rule_inclusion}}). 
% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least for now

@@

% Liz: I know this section is getting long, and maybe this is too blunt, but: 
% People who have experience with exclusion have the social and technical 
% expertise about removing barriers to inclusion. Actively ask where the barriers are and how to mitigate them.
% Rocío: let's make sure this is included somewhere.

The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups. First, rethink power relationships inside the group and share power. Be committed to examine power relations inside your organizing team and challenge it during all the process. % Rocío: Could we explain power relationship / share power ideas in more simple terms?
Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer leading positions, and let people from privileged groups step down. 
Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
% Liz: Use of "voice" in this way makes deaf people feel excluded. Maybe it could be something like "Build an environment in which every person has input..." Or anything as long as it doesn't suggest you have to have a voice to participate.
%ast: i rephrased slightly, avoiding "support" it's not a question of privileged people giving support but of positions being given importance
Offer support and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See \textbf{Rule \ref{rule_language}}), or others. You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership. Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note-taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}). Make work visible, discuss collectively what should be done (and why), divide and account work. Being transparent and sharing the information is key to make work visible and collective. %sara: this relates to make labor visible from DF, maybe we can cite them here too. %Rocío: sure.
% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.

% Liz: I have a little concern about mentoring WRT English, but I'm a white English speaker, so take it with a grain of salt. I think people who look like me need to do more accepting and less mentoring around English. If someone explicitly wants my help, I'm happy to give it because I love words and language, and I might learn something cool about another language. But I'm not sure we should be telling the English-priveleged readers that they are language mentors.
%ast. i agree with liz and wrote this from the point of view of a non-native speaker- in terms of support rather than mentorship. replaced with support but feel free to check
Most importantly, take care of your team. 
Having a diverse team and executing inclusive and accessible practices throughout the organizing period--it might be a year or two--may require a lot of effort. The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone. 
However, having a strong community as the only reward may be enough for those who are in more privileged position. 
Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free. 
% Liz: Yes! This is especially true when *some* people are paid to work on the conference. I was just asked to do accessibility work that should have been done by the contracted, paid organizer. This is hugely problematic and will be true for lots of conferences outside the open software community.
In addition, tasks such as receiving and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
Prioritize your team's well-being. Check on them regularly and make sure everyone is comfortable. 
Be mindful of each particular context, be flexible with hours or commitments, and revise your budget (see \textbf{Rule \ref{rule_financial}}). 
% Rocío: Idea of this paragraph: take care of your team.
% Liz: also, include people from diverse backgrounds in all aspects of planning, not just the diversity-related ones, unless that is their only interest. In some organizations, I could be on the scientific committee and participate in many other ways, but I'm only called on for accessibility.

@@

%idea1: our lists are biased

When choosing or inviting people as keynote speakers, program committee, session chairs, and other spotlight roles, it is likely that there will not be much diversity in the first set of names. Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. Rather than deter us, this implicit and systemic bias should encourage us to look further to find great people that are not routinely in the spotlight. 
Ensuring diversity in each of these roles needs to be a deliberate process. We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.

Make sure that every selection committee--the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees--are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them. 
An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into. 
For example, for useR!2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
When organizing an international academic conference, you may reach local associations, student organizations, or other Early Career Researchers groups or communities of practice: ask for their barriers for participation, and explicitly invite them to contribute to the conference. 
% Rocío: Not everyone knows what a community of practice is. Could we add a definition or use other words? Definition: "Communities of practice are groups of people who share a concern, a set of problems, or a passion about a topic, and who deepen their knowledge and expertise in this area by interacting on an ongoing basis". (Wenger, E., McDermott, R. A., & Snyder, W. (2002). Cultivating Communities of Practice: A Guide to Managing Knowledge. Harvard Business Press.)
Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged. Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers and people from minoritized groups, fostering collaboration and building their skills.
% Liz: maybe consider a broader group than "early career researchers." One problem is that marginalized people have many fewer opportunities, and as they age that attenuation of opportunities compounds. Someone can be in the margins even years after finishing grad school or being a postdoc (not to mention that people without degrees could have contributions.

Furthermore, consider bringing to the spotlight the people that have contributed to your field in more collaborative ways \cite{cheng2020x+}. You can, for instance, reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness. Community building, for instance, is challenging and usually unrewarded when compared to publications or software development \cite{acionWhyChooseCommunity2020}; give the people who deliberately took the time to work on their communities the recognition they deserve. Defy the stereotypical criteria for success by acknowledging these community practices. 
Finally, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to Diversity, Inclusion, and Accessibility. Recognize their areas of expertise and respect their will regarding participation (or not) in community-building events.
% Rocío: Since the system usually rewards the most competitive people, appealing cvs and so on, I think it's worth it to explain that rewarding collaborative work is also part of unbiasing; not only because it's mostly done by women, people of color, etc., but because this kind of amazing work is usually neglected in our conferences. It may not be obvious to everyone that the content of this paragraph is not only "bring these people to the spotlight", but that it also means unbiasing the conference. I don't think I'll have time to add lines about this in the beginning of the paragraph so if anybody could do that, it would be great.

% Liz: I might be able to, but I'm not sure where you want the additional text; where the beginning of the paragraph is because of how this is formatted. I couldn't agree more that being truly inclusive means not restricting people's role. Every subteam should be diverse, in addition to the whole organizing team. It's not just about representation but needs to encompass integration across the whole event.

% ast: i agree a lot with Rocío here, the change in what is given importance is a key part of the unbiasing. i had the impression that the part of not restricting people's role goes above, in team rule.

@@

Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}. Even when the organizing team implements other inclusive practices, the concern about the participation of people with disabilities tends to be partial, and is one of the .

@@

If the conference is in person, the venue must be accessible for people who use wheelchairs, have important signs in Braille, and a sound system compatible with hearing devices, just to name a few important features. Presenters should always speak into a microphone. Take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, menus that include sugar-free, gluten-free options. 
%ast: I removed the captioners part because it goes slightly off-topic and the sound system is apart for them when in person. 
Regardless of the conference format --online, in person, or hybrid-- all images used in the communication strategy of the conference should have alternative text, including the website. Likewise, any videos should have both captions and sound.

Platforms for conference registration and abstract submission, websites, and chat platforms --if used-- should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection). 
Captioning or interpretation for presentations should be available in more than one language if possible. Sign languages are more complicated, because the vary from region to region, but if your conference has a clearly regional scope, a specific sign language (such as ASL, American Sign Language, or LIBRAS, the Brazilian Sign Language) can be adopted as official.

Do not forget about making your social events accessible and making sure that the materials presented are accessible for all. This includes physical activities that should not restrict participation based on body type or ability.

>>>>

% Rocío: Idea of this paragraph: build a diverse team
A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team.
Build a team with people from different regions, genders, races and ethnicity, socioeconomic statuses, and other aspects of diversity. 
Go beyond balancing all genders in this effort and pay special attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}). 
To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. 
If you already have an organizing team, check for gaps in its composition.

% Rocío: Idea of this paragraph: real inclusion when having diverse members
Gathering a diverse team will only work if there is real inclusion. 
People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity. 
This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups \cite{costanzachockDesign2020}.
People who have experience with exclusion have the social and technical expertise about removing barriers to inclusion.
A truly inclusive and welcoming space is one in which everyone in the team is invited to be part of the solution.
However, do not expect people from minoritized groups to only work on these solutions, and respect their interests in other areas of the organization.

% Rocío: Idea of this paragraph: positive outcomes of having a diverse team - keeping it general at least
Creating and maintaining such a team and space may seem more challenging than working in homogeneous teams--when you are not used to them--but the positive outcomes are worthwhile. 
Having diverse people in decision-making positions will affect positively all the other aspects of your conference because all the processes will benefit from their input, expertise, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
In addition, a diverse team plays an important role at creating a welcoming space because representation--seeing people with similar life experiences occupy public spaces, positions of power, and breaking negative stereotypes--is one of the best ways to create a sense of belonging (see \textbf{Rule \ref{rule_inclusion}}).

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.
The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups. 
Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly, offer leading positions, and let people from privileged groups step down. 
Build an environment in which every person can express their position and give priority to people from systematically excluded groups.
Offer support and encouragement if you encounter cases of impostor syndrome. 
You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership.

% Rocío: Idea of this paragraph: take care of your team.
Most importantly, take care of your team. 
Having a diverse team and executing inclusive and accessible practices throughout the organizing period--it might be a year or two--may require a lot of effort. 
The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone. 
However, having a strong community as the only reward may be enough for those who are in more privileged positions. 
Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free; consider paying them as an item in your budget.  
In addition, tasks such as receiving and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded.
Prioritize your team's well-being. Check on them regularly and make sure everyone is comfortable. 
Be mindful of each particular context, be flexible with hours or commitments, and revise your budget (see \textbf{Rule \ref{rule_financial}}).

% % Rocío: might go to the next rule
% Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note-taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}). Make work visible, discuss collectively what should be done (and why), divide and account work. Being transparent and sharing the information is key to make work visible and collective. %sara: this relates to make labor visible from DF, maybe we can cite them here too. %Rocío: sure.

@@

@@

%idea of paragraph 1: our lists are biased
When choosing or inviting people as keynote speakers, program committee, session chairs, and other spotlight roles, it is likely that there will not be much diversity in the first set of names. 
Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. 
Rather than deter us, this implicit and systemic bias should encourage us to look further to find great people that are not routinely in the spotlight. 
Ensuring diversity in each of these roles needs to be a deliberate process. 
We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions. 
Make sure that every selection committee--the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees--are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them. 
An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into.

% Rocío: main idea: unbias roles and give spotlight to the roles--and people who have contributed in roles--that are not stereotypically categorized as success
In academia and the industry, not only does the list of people considered in spotlight roles is usually biased, but the spotlight roles themselves are.
Scientific publication and development of software, are often regarded with higher value than community building or teaching, which may be equally if not more challenging, less rewarded, and usually tackled by women, people of color, people with disabilities and from other minoritized groups \cite{cheng2020x+}.
Defy the stereotypical criteria for success by acknowledging these community practices and the people behind them.
You can, for instance, give more space in your conference to talks about community building, and reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness. 
Finally, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to diversity and inclusion; recognize their areas of expertise and interest.

@@

Conferences are among the least accessible spaces that people with disabilities may encounter \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. This is one key aspect where having someone in the organizing team allows them to take part in the decisions from the beginning (see \textbf{Rule \ref{rule_organizing_team}}), as thinking about disability or simulating it is not a substitute for real-life experience \cite{costanzachockDesign2020}. Accessibility requires time and early decision-making \cite{irishIncreasingParticipationUsing2020} and should not be dealt with at the last minute. Conversely, inaccessible decisions are hard to correct, e.g., when finding out too late that a venue is inaccessible for people who use wheelchairs.

@@

If the conference is in person, the venue must be accessible for people who use wheelchairs, have important signs in Braille, and a sound system compatible with hearing devices, just to name a few important features. Presenters should always speak into a microphone. Take care proactively of invisible disabilities, for example, by providing quiet spaces for privacy and noise-free conversations, menus that include sugar-free, gluten-free options.

Regardless of the conference format --online, in person, or hybrid-- all images used in the communication strategy of the conference should have alternative text, including the website. Likewise, any videos should have both captions and sound.

Platforms for conference registration and abstract submission, websites, and chat platforms --if used-- should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection). 
Captioning or interpretation for presentations should be available in more than one language if possible. 
Do not forget about making your social events accessible and making sure that the materials presented are accessible for all. This includes physical activities that should not restrict participation based on body type or ability.
